---
title: "Hypertuning on IMDB dataset"
output:
  word_document: default
  html_notebook:
    highlight: textmate
    theme: cerulean
  html_document:
    df_print: paged
---
# The IMDB dataset

Load IMDB dataset
The IMDB dataset is preloaded in the Keras

```{r}
library(keras)
library(tensorflow)
library(ggplot2)
library(tidyverse)
library(cowplot)
imdb <- dataset_imdb(num_words = 10000)
c(c(train_data, train_labels), c(test_data, test_labels)) %<-% imdb
```

The argument num_words = 10000 means that we will only keep the top 10,000 most frequently occurring words in the training data. Rare words will be discarded

The variables train_data and test_data are lists of reviews, each review being a list of word indices (encoding a sequence of words). train_labels and test_labels are lists of 0s and 1s, where 0 stands for “negative” and 1 stands for “positive”:

# Preparing the Data

```{r}
vectorize_sequences <- function(sequences, dimension = 10000) {
  # Create an all-zero matrix of shape (len(sequences), dimension)
  results <- matrix(0, nrow = length(sequences), ncol = dimension)
  for (i in 1:length(sequences))
    # Sets specific indices of results[i] to 1s
    results[i, sequences[[i]]] <- 1
  results
}

# Our vectorized training data
x_train <- vectorize_sequences(train_data)
# Our vectorized test data
x_test <- vectorize_sequences(test_data)
# Our vectorized labels
y_train <- as.numeric(train_labels)
y_test <- as.numeric(test_labels)
set.seed(123)
val_indices<-sample(1:nrow(x_train),nrow(x_train*(0.40)))
x_val <- x_train[val_indices,]
partial_x_train <- x_train[-val_indices,]

y_val <- y_train[val_indices]
partial_y_train <- y_train[-val_indices]

```

# Building our network and Validating our Approach

Single layer network without any technique

```{r}
model <- keras_model_sequential() %>% 
  layer_dense(units = 16, activation = "relu", input_shape = c(10000)) %>%
  layer_dense(units = 1, activation = "sigmoid")


model %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("accuracy")
)

history <- model %>% fit(
  partial_x_train,
  partial_y_train,
  epochs = 20,
  batch_size = 512,
  validation_data = list(x_val, y_val)
)

str(history)
plot(history)

model %>% fit(x_train, y_train, epochs = 4, batch_size = 512)
results <- model %>% evaluate(x_test, y_test)
results
```

Single layer network using regularization, 'mse' loss function and tanh activation

```{r}
model <- keras_model_sequential() %>% 
  layer_dense(units = 64,  kernel_regularizer = regularizer_l2(.0001),activation = "tanh", input_shape = c(10000)) %>%
  layer_dense(units = 1, activation = "sigmoid")

model %>% compile(
  optimizer = "rmsprop",
  loss = "mse",
  metrics = c("accuracy")
)

history <- model %>% fit(
  partial_x_train,
  partial_y_train,
  epochs = 20,
  batch_size = 512,
  validation_data = list(x_val, y_val)
)

str(history)
plot(history)

model %>% fit(x_train, y_train, epochs = 5, batch_size = 512)
results <- model %>% evaluate(x_test, y_test)
results
```

Single layer network with regularization and drop out

```{r}
model <- keras_model_sequential() %>% 
  layer_dense(units = 16,  kernel_regularizer = regularizer_l2(.0001),activation = "tanh", input_shape = c(10000)) %>%
  layer_dropout(rate = 0.5) %>%
  layer_dense(units = 1, activation = "sigmoid")

model %>% compile(
  optimizer = "rmsprop",
  loss = "mse",
  metrics = c("accuracy")
)

history <- model %>% fit(
  partial_x_train,
  partial_y_train,
  epochs = 20,
  batch_size = 512,
  validation_data = list(x_val, y_val)
)

str(history)
plot(history)

model %>% fit(x_train, y_train, epochs = 7, batch_size = 512)
results <- model %>% evaluate(x_test, y_test)
results
```

Triple layer network without using regualrization, setting units to 32 and dropout

```{r}
model <- keras_model_sequential() %>% 
  layer_dense(units = 32, activation = "tanh", input_shape = c(10000)) %>%
  layer_dense(units = 32, activation = "tanh") %>%
  layer_dense(units = 32, activation = "tanh") %>%
  layer_dense(units = 1, activation = "sigmoid")

model %>% compile(
  optimizer = "rmsprop",
  loss = "mse",
  metrics = c("accuracy")
)

history <- model %>% fit(
  partial_x_train,
  partial_y_train,
  epochs = 20,
  batch_size = 512,
  validation_data = list(x_val, y_val)
)

str(history)
plot(history)

model %>% fit(x_train, y_train, epochs = 4, batch_size = 512)
results <- model %>% evaluate(x_test, y_test)
results
```

Triple layer network by regualrization

```{r}
model <- keras_model_sequential() %>% 
  layer_dense(units = 64, kernel_regularizer = regularizer_l2(.0001), activation = "tanh", input_shape = c(10000)) %>%
  layer_dense(units = 32, kernel_regularizer = regularizer_l2(.0001), activation = "tanh") %>%
  layer_dense(units = 32, kernel_regularizer = regularizer_l2(.0001), activation = "tanh") %>%
  layer_dense(units = 1, activation = "sigmoid")

model %>% compile(
  optimizer = "rmsprop",
  loss = "mse",
  metrics = c("accuracy")
)

history <- model %>% fit(
  partial_x_train,
  partial_y_train,
  epochs = 20,
  batch_size = 512,
  validation_data = list(x_val, y_val)
)

str(history)
plot(history)

model %>% fit(x_train, y_train, epochs = 2, batch_size = 512)
results <- model %>% evaluate(x_test, y_test)
results
```

Triple layer network by regualrization and dropout

```{r}
model <- keras_model_sequential() %>% 
  layer_dense(units = 64, kernel_regularizer = regularizer_l2(.0001), activation = "tanh", input_shape = c(10000)) %>%
  layer_dropout(rate = 0.5) %>%
  layer_dense(units = 32, kernel_regularizer = regularizer_l2(.0001), activation = "tanh") %>%
  layer_dropout(rate = 0.5) %>%
  layer_dense(units = 32, kernel_regularizer = regularizer_l2(.0001), activation = "tanh") %>%
  layer_dropout(rate = 0.5) %>%
  layer_dense(units = 1, activation = "sigmoid")

model %>% compile(
  optimizer = "rmsprop",
  loss = "mse",
  metrics = c("accuracy")
)

history <- model %>% fit(
  partial_x_train,
  partial_y_train,
  epochs = 20,
  batch_size = 512,
  validation_data = list(x_val, y_val)
)

str(history)
plot(history)

model %>% fit(x_train, y_train, epochs = 2, batch_size = 512)
results <- model %>% evaluate(x_test, y_test)
results
```


Using epochs, we train new network from scratch. Later evaluate on the test data.

```{r}
model1 <- keras_model_sequential() %>% 
  layer_dense(units = 64, kernel_regularizer = regularizer_l2(.0001), activation = "tanh", input_shape = c(10000)) %>%
  layer_dropout(rate = 0.5) %>%
  layer_dense(units = 64, kernel_regularizer = regularizer_l2(.0001), activation = "tanh") %>%
  layer_dropout(rate = 0.5) %>%
  layer_dense(units = 64, kernel_regularizer = regularizer_l2(.0001), activation = "tanh") %>%
  layer_dropout(rate = 0.5) %>%
  layer_dense(units = 1, activation = "sigmoid")

model1 %>% compile(
  optimizer = "rmsprop",
  loss = "mse",
  metrics = c("accuracy")
)
history1<-model1 %>% fit(
  x_train,
  y_train,
  epochs = 7,
  batch_size = 250)

modelb.df <- as.data.frame(history1$metrics)
modelb.df <- modelb.df %>% mutate(epochs=1:n())
p11<-ggplot(modelb.df) + geom_line(aes(x=epochs,y=loss),size=0.7,color="blue")+geom_point(aes(x=epochs,y=loss),size=1,color="blue")+ggtitle("Epochs vs Loss function ")+theme(panel.grid = element_blank(),panel.background = element_blank())+theme_classic()+geom_text(aes(epochs,loss,label=round(loss,2)),vjust=-0.2)
p12<-ggplot(modelb.df) + geom_line(aes(x=epochs,y=accuracy),size=0.7,color="blue")+geom_point(aes(x=epochs,y=accuracy),size=1,color="blue")+ggtitle("Epochs vs Accuracy")+theme(panel.grid = element_blank(),panel.background = element_blank())+theme_classic()+geom_text(aes(epochs,accuracy,label=round(accuracy,2)),vjust=-0.2)
plot_grid(p11,p12,nrow = 2)
results<- model1 %>% evaluate(x_test,y_test)
results
```

